---
title: "Lab 03 - Nobel laureates"
author: "Ryan Wheat"
date: "01/28/23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are 26 variables, and 935 observations within the Nobel dataset. Each row represents one nobel prize winner, and their subsequent information.

```{r glimpse data}
glimpse(nobel)
```

### Exercise 2

I have 228 observations!

```{r filters}

#creating new dataset

nobel_living <- nobel %>%
  filter(!is.na(x = country)) %>%
  filter(gender != "org") %>%
  filter(is.na(x = died_date))

#making sure i did it right
glimpse(nobel_living)
```


```{r us based winners}

nobel_living <- nobel_living %>%
  mutate(country_us = if_else(country == "USA", "USA", "Other"))

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```


### Exercise 3

It does seem as though, across all categories, nobel prize winners tended to be located in the United States when receiving their award. This provides support the article's claims so far.

```{r location visuals}

#bar graph plot
ggplot(nobel_living_science, aes(x = country_us)) +
  geom_bar() +
  facet_wrap(~category) +
  coord_flip()
```

### Exercise 4

105 out of 228 nobel prize winners were born in the U.S.

```{r birth location}

#creating new birth variable

nobel_living_science <- nobel_living_science %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other"))

#count of usa vs. other birth countries
nobel_living_science %>%
  count(born_country_us)
```


### Exercise 5

The majority of US-based nobel prize winners were also born in the US -- but a noticeable portion were not for each prize category. This seems to support BuzzFeeds claims that immigration is important for scientific pursuits.

```{r nobel winner current location and birth}

#bar graph plot from before but adding fill
ggplot(nobel_living_science, aes(x = country_us, fill = born_country_us)) +
  geom_bar() +
  facet_wrap(~category) +
  coord_flip()
```


### Exercise 6

UK and Germany are tied for the most common at 7 nobel prize winners who were based in the US.

```{r is uk best outside country?}

#filter out USA born from USA-based winners
nobel_living_science %>%
  filter(country_us == "USA" & born_country_us != "USA") %>%
  #count the countries remaining
  count(born_country) %>%
  arrange(desc(n))

```

